---
title: "EverythingYouWillNeed.com Sales Anomalies and Forecast Report"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: 
      version: 3
      primary: "#008080"
      navbar-bg: "#008080"
      base_font: 
        google: Prompt
      heading_font:
        google: Sen
      code_font:
        google: 
          family: JetBrains Mono
          local: false
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(shinyWidgets)
library(shinyjs)
library(data.table)
library(tidyverse)
library(tidyquant)
library(plotly)
library(anomalize)
library(modeltime)
library(timetk)
library(tidymodels)

```

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>

```{r}

product_purchases <- read.csv("product_purchase_overtime.csv")

first_purchase_date <- product_purchases %>%
  mutate(Date = as.Date(Date)) %>% 
  filter(purchase_count != 0) %>%
  arrange(products, Date) %>% 
  group_by(products) %>% 
  mutate(mindate = min(Date)) %>% 
  distinct(products, mindate) %>% 
  ungroup()

product_purchases <- product_purchases %>% 
  mutate(Date = as.Date(Date)) %>% 
  as_tibble() %>% 
  select(products, Date, purchase_count) %>% 
  left_join(first_purchase_date) %>% 
  filter(Date >= mindate) %>% 
  select(-mindate)

```


Sidebar {.sidebar}
=====================================

```{r}
# Inputs

selectizeInput(
    inputId = "input_product",
    label = h5("Select Your Product and click on Apply to get started"),
    choices = unique(product_purchases$products),
    selected = "Silent Canary alarm",
    multiple = TRUE,
    options = list(maxItems = 1)
)

hr()

actionButton(inputId = "apply", label = "Apply", icon = icon("play"),  style="color: #fff; background-color: #008080; border-color: #008080;")

actionButton(inputId = "reset", label = "Reset", icon = icon("sync"),  style="color: #fff; background-color: #008080; border-color: #008080")
```



```{r}
hr()

dateRangeInput(
    inputId = "date_range",
    label = h5("Input Date Range"),
    start = min(product_purchases$Date),
    end = max(product_purchases$Date),
    min = min(product_purchases$Date),
    max = max(product_purchases$Date),
)

hr()

selectInput(
        inputId  = "input_alpha",
        label    = h5("Alpha (Sensitivity for IQR anomaly detection)"),
        choices  = seq(0.00, 1, by=0.01),
        multiple = FALSE,
        selected = 0.05
)

hr()
```



```{r}


observeEvent(eventExpr = input$reset, handlerExpr = {
  
  updateSelectizeInput(
    session = session,
    inputId = "input_product",
    selected = "Silent Canary alarm"
    )
  
  updateDateRangeInput(
    session=session,
    inputId = "date_range",
    start = min(product_purchases$Date),
    end = max(product_purchases$Date)
  )
  
  updateSelectInput(
    session = session,
    inputId  = "input_alpha",
    selected = 0.05
    
  )
  
})
```


<hr>  
<br>
<br>

___Note: This report is based on fake data and fake company name. All product names are made up. Any resemblance to anything fictional or otherwise is purely coincidental___

<hr>


```{r}
products_distinct <- product_purchases %>% distinct(products)

products_count_filtered <- eventReactive(
  eventExpr = input$apply,

  valueExpr = {
      product_purchases %>%
          filter(products %in% input$input_product) %>% 
          filter(Date %>% between(left  = input$date_range[1],
                                           right = input$date_range[2]))
  },
  ignoreNULL = FALSE
)


product_decomposed_iqr <- eventReactive(
  eventExpr = input$apply,

  valueExpr = {
      products_count_filtered() %>%
            time_decompose(purchase_count, merge = TRUE) %>%
            anomalize(target = remainder,
                      method = "iqr", 
                      alpha  = as.double(input$input_alpha)) %>%
                      mutate(anomaly = ifelse(purchase_count <= 0, "Yes", anomaly)) %>%
                      group_by(products, group = cumsum(anomaly != lag(anomaly, default = first(anomaly)))) %>%
                      mutate(counter = row_number()) %>%
                      ungroup() %>%
                      select(-group) %>%
                      mutate(counter = as.double(counter)) %>%
                      time_recompose()  %>%
                      mutate(purchase_count_cleaned = ifelse(anomaly == "Yes", season + trend, observed))
  },
  ignoreNULL = FALSE
)

product_ts_tbl <- eventReactive(
  eventExpr = input$apply,

  valueExpr = {
      products_count_filtered() %>%
      select(Date, purchase_count ) %>%
      set_names(c("date", "value"))
  },
  ignoreNULL = FALSE
)
```

```{r}
calibration_forecast_tbl <- eventReactive(
  eventExpr = input$apply,

  valueExpr = {
      splits <- product_ts_tbl() %>%
      time_series_split(assess = "1 month", cumulative = TRUE)
    
    model_fit_prophet <- prophet_reg() %>%
      set_engine("prophet", yearly.seasonality = FALSE) %>%
      fit(value ~ date, training(splits))
    
    model_fit_prophet %>%
      modeltime_calibrate(testing(splits))
  },
  ignoreNULL = FALSE
)


```




Product Sales
=====================================

Column
-----------------------

### 

```{r}

output$products_ts <- renderPlotly({
    g <- product_ts_tbl() %>%
      plot_time_series(date, value, .interactive = TRUE, .smooth = FALSE, .title = str_glue("Sales over time")) 
    ggplotly(g)
    
})

plotlyOutput(outputId = "products_ts")

```


Column {data-width=350}
-----------------------

### Total purchase count for filtered data

```{r}

renderValueBox({
  valueBox(
    value   = products_count_filtered() %>%
      summarize(total_purchase = scales::number(sum(purchase_count, na.rm=TRUE), big.mark=",")) %>% 
      pull(total_purchase),
    caption = "Total product sales",
    color   = "rgba(0,128,128,0.2)"
    )
})


```


### Avgerage weekly purchase 

```{r}

renderValueBox({

  valueBox(
    value   = products_count_filtered() %>% 
  mutate(weeklydate = FLOOR_WEEK(Date)) %>% 
  group_by(products, weeklydate) %>%
  summarise(weeklypurchase = sum(purchase_count, na.rm=TRUE)) %>% 
  ungroup() %>% 
  summarise(scales::number(mean(weeklypurchase, na.rm=TRUE), big.mark=",")) %>% 
  pull(),
    caption = "Avg weekly sales",
    color   = "rgba(0,128,128,0.2)"
    )

})

```

### Products Purchase table filtered

```{r}
DT::renderDataTable(products_count_filtered() %>% arrange(desc(Date)),
                          extensions = 'Buttons',
                          options = list(
                                scrollX = TRUE,
                                scrollY = "250px",
                                pageLength = 100,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            )
  )
```






<!-- > This dashboard looks at data from _EverythingYouWillNeed.com_ website to report out on daily product sales. You can select the product, the dates to analyze and click on apply to replicate the report for the specific product choice and time period. -->




Forecast Sales {data-orientation=rows}
=====================================

Row {data-height=980}
-------------------------------------


### Forecast plot

```{r eval=TRUE}

output$product_forecast_plot <- renderPlotly({
    showModal(modalDialog("Calculating forecast", footer=NULL))
    g <- calibration_forecast_tbl() %>%
    modeltime_refit(product_ts_tbl() %>% filter(value !=0)) %>%
      modeltime_forecast(h = "1 month", actual_data = product_ts_tbl()) %>%
      plot_modeltime_forecast(.interactive = TRUE, .conf_interval_show = TRUE, .smooth = FALSE, .legend_show = TRUE, .color_lab = "Forcast method")
  removeModal()
  ggplotly(g)
})

plotlyOutput(outputId = "product_forecast_plot")

```


Row 
-------------------------------------

### Forecast output

```{r eval=TRUE}

DT::renderDataTable(as.data.frame(calibration_forecast_tbl() %>%
    modeltime_refit(product_ts_tbl() %>% filter(value !=0)) %>%
      modeltime_forecast(h = "1 month", actual_data = product_ts_tbl()) %>% 
      arrange(desc(.index))),
                          extensions = 'Buttons',
                          options = list(
                                scrollX = TRUE,
                                scrollY = "250px",
                                pageLength = 100,
                                dom = 'tB',
                                buttons = c('copy', 'csv', 'excel')
                            )
  )

```

Anomaly Detection {data-orientation=columns}
=====================================

Columns
-----------------------

### Anomalies plot

> Change alpha on the sidebar and click on _"Apply"_ to increase or decrease the threshold of what is considered anomalous data

```{r}
output$product_anomalies <- renderPlotly({
    showModal(modalDialog("Calculating anomalies", footer=NULL))
    g <- product_decomposed_iqr() %>%
            plot_anomalies(time_recomposed = TRUE) +
            expand_limits(y = 0) +
            scale_y_continuous(labels = scales::number_format(big.mark = ",")) +
            labs(x = "", y = "Count") +
            geom_line(aes(Date, purchase_count), color = "darkgreen") +
            expand_limits(y = 0) +
            scale_y_continuous(labels = scales::number_format(big.mark = ","))  +
            theme_tq() + theme(legend.position = "none") +
            theme(
              axis.text.x = element_blank()
            )
  removeModal()
  ggplotly(g)
})

plotlyOutput(outputId = "product_anomalies") 
```


About this dashboard
=====================================

Column
-----------------------

### About this dashboard 

#### This dashboard looks at data from _EverythingYouWillNeed.com_ website to provide insights on daily product sales by allowing users to select a product, a time period, and see the corresponding sales data. The dashboard also includes a forecast tab and an anomalies tab that helps users to forecast future sales and identify any anomalies in upstream data.

### About Product Sales tab

#### The Product Sales tab provides an overview of the product sales and key performance indicators (KPIs) over time. It includes a table that shows the total sales of the selected product over time. The table and plot can be filtered based on the selected product and time period. Package used for this plot is  [timetk](https://business-science.github.io/timetk/). 

### About Forecast Sales tab

#### The Forecast tab helps users to forecast future sales for the selected product. It includes a forecast plot that shows the predicted sales for the next month, and a table that provides the details of the forecast. Packages used for this plot are [modeltime](https://business-science.github.io/modeltime/), and [tidymodels](https://www.tidymodels.org/). The table can be downloaded in different formats like csv, and excel. 

### About Anomalies tab


#### The Anomalies tab checks for anomalies in the sales data for the selected product. It includes a time series plot that shows the sales data with anomalies highlighted. Package used for this plot is [anomalize](https://business-science.github.io/anomalize/). It can also be used to create a decomposed time series plot that shows the seasonality, trend, and remainder of the time series data. The alpha parameter can be changed to increase or decrease the threshold of what is considered anomalous data. 

